Nature or Nurture? Collisionless Evolution of Galactic Disc-Halo Systems

### 2019-09-18 Thesis/Paper_III.tex
More updates. On to final minutia

\chapter{Forming Vertical Structure in $\Lambda$CDM Discs}\label{ch:paper_iii}

>>>>

\newcommand{\kpch}{\ensuremath{h^{-1} \, \text{kpc}}}
\newcommand{\kpc}{\ensuremath{\text{kpc}}}
\newcommand{\solarmh}{\ensuremath{h^{-1} \, M_\odot}}
\newcommand{\solarm}{M_\odot}
\chapter{Forming Vertical Structure in $\Lambda$CDM Discs}\label{ch:paper_iii}
\textbf{This chapter contains a draft of a paper in preparation to be submitted to Monthly Notices of the Royal Astronomical Society. The working title for the paper is ``Vertical Structure and Kicked-Up Stars in $\Lambda$CDM Galaxies". }

\newpage
\section{Abstract}

We perform numerical experiments where stellar discs are inserted into
dark matter halos in a $\Lambda$CDM simulation after which the
fully-live disc-halo systems are evolved until the present epoch. {Prior it its insertion, a disc is treated as a growing, rigid potential whose position and orientation are determined from rigid-body dynamics. When the disc goes live, the disc-halo system is in approximate equilibrium.} The
discs develop structures such as bars, warps, and bending waves.  Not
surprisingly, we find that when we insert the disc at a redshift
$z=1$, it develops more structure and exhibits more clearly signs of
disequilibrium then when it is inserted later.  We identify
populations of stars kicked out of the disc to high latitudes that are
qualitatively similar what is observed in the Milky Way. {An in-depth attempt is made to disentangle the effects of bar buckling, tidal fields due to 
the generally triaxial halo and from distant subhalos, and the
passage of subhaloes through the inner disc. }While our
results are consistent with the hypothesis the the kicked-out-disc was
produced by a very massive satellite galaxy, we also find that they
can arise from the interaction of the disc with its halo. We conclude
that in the Milky Way a wide variety of cosmological environments are
able to explain disc structures like Monoceros, A13, and TriAnd.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:
\section{Introduction} \label{sec:introduction}

The \textsc{Hi} discs of late-type galaxies, when observed edge-on,
often exhibit pronounced warps (see, for example,
\citet{sancisi_1976,bosma_1991, garcia-ruiz_2002} and reviews by
\citet{binney_1992} and \citet{Sellwood2013}). Indeed, the \textsc{Hi}
warp in the Milky Way has been known since the work of
\citet{oort_1958}. There is also evidence for a stellar warp in the
Milky Way, which roughly traces that of the \textsc{Hi} disc
\citep{cox_1996, reyle_2009}.

The prevalence of warps suggests that a typical stellar disc
experiences perturbations normal to its midplane. Recent surveys,
particularly of the stellar content in the Galaxy, have revealed other
manifestations of vertical perturbations. For example, surveys such as
the Sloan Extension for Galactic Understanding and Exploration (SEGUE)
and Gaia Data Release 2 (GDR2) have uncovered an asymmetry with
respect to the Galactic midplane in stellar number counts in the
vicinity of the Sun \citep{widrow_2012_sdss, yanny_gardner_2013,
  bennet_2019_gaia}. In addition, bulk motions of stars normal to the
midplane have been found in data from SEGUE, the Radial Velocity
Experiment (RAVE), the LAMOST survey, and GDR2
\citep{widrow_2012_sdss, williams_2013_rave, carlin_2013_lamost,
  pearl_2017, carrillo_2018_rave, gaia_collab}. Furthermore, a
corrugation in stellar number counts with a wavelength of $\sim
5\,{\rm kpc}$ and an amplitude of $\sim 100\,{\rm pc}$ has been
detected between $10$ and $15$ kpc in Galactocentric radii
\citep{xu_2015}. Corrugations in the velocity field of gas discs
in external galaxies have also been observed \citep[for
  example]{matthews_2008,sanchez_2015}.

In addition to waves and corrugations, several stream-like stellar
structures in the Milky Way may have originated within the
disc. In the past, these structures were thought to comprise stars
tidally stripped from disrupted dwarf galaxies or globular
clusters. Recently, the idea that some of the streams are actually
stars 'kicked out' of the disc has gained traction. One example is the
Monoceros ring \citep{newberg_2002, yanny_2003}, also called the
Galactic Anticentre Stellar Structure (GASS) \citep{crane_2003,
  rocha-pinto_2003}, which is located at low latitudes towards the
Galactic anti-centre at a Galactocentric distance of about
$18\,\,\kpc$. The coincidence in position and velocity of GASS and
disc stars suggests that a disc origin may be a more likely explanation than tidal stripping \citep{deason_2018,
  monoceros_disk_origin}. Similar structures have been found in the
outer disc, most notably the Triangulum-Andromeda clouds (TriAnd)
\citep{triand_discovery} and A13 \citep{a13_discovery}. There is also
evidence for kicked-out stars in M31 \citep{richardson_2008_m31,
  dorman_2013_m31, bernard_2015_m31} and M33
\citep{mcconnachie_2006_m33, mcconnachie_2010_m33}. (For a review of
the kicked-out-disc hypothesis, see \citet{johnston_kud_review}.)

There is now a concerted effort to explain vertical disequilibria in
Milky Way-like galaxies. Much of this effort has focused on the
hypothesis that a single massive subhalo, such as the Sagittarius
dwarf spheroidal galaxy (Sgr dSph) \citep{ibata_discovery}, is
responsible for much of the vertical structure in the Milky
Way. Simulations of a single satellite encounter with an isolated
galaxy have been used to study this hypothesis \citep[for
  example]{purcell2011,gomez_2013,widrow_2014,feldmann_2015, dlv_2015,
  donghia_2016, laporte_2016, laporte_2018, laporte_2018_b}. While
many of the observed features are easily reproduced in the
simulations, it appears that a very massive ($10^{11}\,\solarm$) Sgr
dSph-like satellite is required to explain the extraplanar streams in
the outer disc \citep{laporte_2018_b}.

The single-satellite encounter is a rather idealized scenario;
disc-environment interactions in real galaxies will, of course, be
more complicated. In particular, discs may be continually perturbed by
any number of satellites or dark matter subhaloes whose orbits take
them into the disc region of the galaxy. Early steps toward a more
realistic model were taken by \citet{font_2001} and
\citet{gauthier_2006} who allowed for a system of subhalos. In
particular, \citet{gauthier_2006} simulated an M31-like disc in a dark
halo where roughly 10\% of the smooth halo mass was replace by dark
subhalos with a mass spectrum motivated by cosmological $\Lambda$CDM
simulations. They found that one or more of the subhalos triggered the
formation of a bar. Further analyses of this simulation and a similar
one by \citet{chequers_2018} found that the system of subhalos could
excite a plethora of vertical structure. However, these simulations
failed to create the streams of disc stars described above.

When a satellite passes through the plane of disc, it interacts with
the disc through a process akin to dynamical friction. Essentially,
disc stars fall in behind the satellite, disrupting and ultimately
heating the disc \citep{sellwood_1998}. On the other hand a distant
but very massive satellite can excite bending modes via differential
acceleration. In the simple example of a satellite falling toward the
disc along its spin axis, the acceleration of the disc toward the
satellite decreases with radius. Thus, as the satellite passes
through the centre of the disc, it will set up axisymmetric bending
modes as seen in \citet{sellwood_1996}.

Finally, we note that dark halos in $\Lambda$CDM cosmologies are
thought to be aspherical and will therefore exert torques on the disc
that can cause it to bend and warp. The effect is due to a
misalignment of the disc symmetry axis with any of the approximate
symmetry axes of the halo. For example, \citet{hu_2016} embed discs
in triaxial haloes. In their appendix, strong warps can be seen in
discs which are substantially misaligned from their
haloes. \citet{gomez_2017} found that vertical structure was
ubiquitous in their hydrodynamical cosmological simulations, some of
which did not have discs in a stable orientation to their host halo.

In this paper, we study the vertical structure that develops in discs
embedded in realistic $\Lambda$CDM haloes. We use a disc insertion
scheme to set up initial conditions where a stellar disc is in
approximate equilibrium with a cosmological halo
\citep{berentzen_2006, debuhr_2012, ys_2015, bauer2018a, hu_2018}. Our
scheme affords us more control over the structural properties of the
disc than would be possible in fully self-consistent hydrodynamical
simulations and was successfully used to study bar formation in a
cosmological context in \citet{bauer2018b}. In fact, bar buckling is
another mechanism for exciting vertical structure in stellar discs
\citep{bar_buckling_echo}, a point we return to later in this paper.

Our paper is organized as follows: In \S\ref{sec:overview} we present
two toy models that serve to illustrate the mechanisms by which
satellites can excite bending modes in a stellar disc. An overview of
the cosmological simulations is given in \S\ref{sec:description} while
the time evolution and key structural features of the discs in these
simulations are discussed in \S \ref{sec:evolution}. In \S\ref{sec:kud}
we examine how stars are kicked out of discs.  We relate our key
results to other studies in \S\ref{sec:discussion} and present a
summary and some concluding remarks in \S\ref{sec:conclusion}.

\begin{figure}
	\centering
	\includegraphics[width=0.85\textwidth]{../figures/isolated_z_0_r_a.eps}
	\caption{Mean vertical displacement, $\langle z \rangle$, of the disc for the
          toy model described in \S\ref{ssec:toy_model_1} as a
          function of $R$ and $t$.} \label{fig:toy_model_1_mean_height}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.85\textwidth]{../figures/isolated_z_0_r_a_halo.eps}
	\caption{Mean vertical displacement of the halo in the toy model described in
          \S\ref{ssec:toy_model_1} as a function of spherical radius $r$ and $t$.} 
        \label{fig:toy_model_1_mean_height_halo}
\end{figure}

\section{Single satellite simulations} \label{sec:overview}

In this section, we illustrate some of the factors driving vertical
structure in stellar discs through two toy models where a disc
interacts with a single subhalo.

\subsection{Axisymmetric subhalo encounter} 
\label{ssec:toy_model_1}

In our first experiment, the subhalo is initially at rest and
positioned $100\,{\rm kpc}$ from the centre of the disc along its
symmetry axis. We generate the initial conditions for the disc, halo,
and subhalo using the \textsc{GalactICS} code
\citep{KGGalactICSReference,WPDGalactICSReference}. The disc has an
exponential surface density with scale length $R_d=3.7\,\kpc$ and mass
$M_d=4\times10^{10} \,\solarm$. The halo has an NFW density profile
\citep{navarro_1997}, with scale length $R_h=33.2\,\kpc$, virial mass
$M_h=1.1\times10^{12}$, and concentration $c=6.7$. We use $10^6$ halo
particles and $3.5 \times 10^{6}$ disc particles. The subhalo is also
assumed to have an NFW profile with scale length $R_h=1.70\,\kpc$,
virial mass $M_h=2.7\times10^{10}$, and concentration $c=38.1$. The
choice of parameters for the disc, halo, and subhalo are motivated by
examples from our suite of cosmological simulations. The simulation is
performed using \textsc{Gadget-3}, a privately released version of the
popular \textsc{Gadget-2} code \citep{GadgetCodePaper}.

In Fig. \ref{fig:toy_model_1_mean_height} we show the azimuthally
averaged vertical displacement, $\langle z\rangle(R,\,t)$. The displacement
is relative to the disc midplane, defined by the mean $z$ of all the
particles. {
This quantifies axisymmetric vertical perturbations, that is, $m=0$
vertical perturbations where $m$ is the azimuthal mode number.  More generally, for a quantity $\Theta$, we define the $m$-fold symmetric $\Theta_m$,}
\begin{eqnarray}
\Theta_{m} &=&\frac{1}{M_S} \int_S \text{d}^3 \textbf{r} \text{d}^3\textbf{v} f(\textbf{r}, \textbf{v}) \Theta(\textbf{r} , \textbf{v})e^{i m \phi_j}\\
&\approx& \frac{1}{M_S} \sum_{j \in S} m_j \Theta(\textbf{r}_j, \textbf{v}_j)e^{im \phi_j} \label{eq:z_statistic}
\end{eqnarray}
{where $M_S$ is the mass in a region $S$, $f(\textbf{r},\textbf{v})$ is the phase space distribution function, $\phi_j$ is the azimuthal angle of the $j$-th particle in $S$, and $m_j$ is the mass of the $j$-th particle in $S$. 
}

At first the disc is relatively unperturbed; the subhalo is
so far away that it exerts an approximately uniform acceleration
across the disc. Only in the very outer disc do we see slight disc
flapping, that is, a lagging of the outer disc relative to the inner
one. The lag is more pronounce as the subhalo makes its first
pericentric passage. After the subhalo has passed through the disc, at
about $1\,{\rm Gyr}$, the pattern reverses as the center of the disc
is pulled back down. The resulting lag of the outer disc then shows up
as a positive signal. The disc develops a corrugation pattern as seen
in the alternating red and blue bands (positive and negative $\langle
z\rangle$). These bending waves propagate outward (bands stretch down
and to the right in the figure.) Over time, the wavelength decreases
as does the speed of propagation. Note that the strong ($> 1\,{\rm
  kpc}$) flapping of the disc beyond $20\,{\rm kpc}$ persists for many
dynamical times (see, also \citet{sellwood_1996}) and weaker
corrugation patters are seen further in.

Qualitatively similar behaviour is seen in the halo. In
Fig. \ref{fig:toy_model_1_mean_height_halo} we show $\langle
z\rangle$, this time averaged in spherical shells in the halo. At
early times, the halo is dragged along with the subhalo. Eventually,
we get a sloshing back and forth of the halo.

\begin{figure}
	\centering
        \includegraphics[width=0.85\textwidth]{../figures/isolated_orbits_two_fits_two_panel.eps}
	\caption{Top panel: Orbit of the disc (blue) and inner halo
          (red) in the toy model described in
          \S\ref{ssec:toy_model_2}. Only the $z$-components of the
          disc and halo position vectors are shown. Bottom panel:
          Orbits of the disc and inner halo (blue) and subhalo
          density peak (red).  The black lines show the fits for the
          two epoch as described in the
          text.} \label{fig:toy_model_2_orbits}
\end{figure}

\subsection{Off-axis subhalo encounter} \label{ssec:toy_model_2}
\begin{figure}
	\centering
	\includegraphics[width=0.85\textwidth]{../figures/b_late_light_subhalo_z_0_r_a.eps}
	\caption{Mean height, $\langle z \rangle$, of the disc for the
          toy model described in \S\ref{ssec:toy_model_2} as a
          function of $R$ and
          $t$.} \label{fig:toy_model_2_mean_height}
\end{figure}

In our second experiment, the satellite is again initialized at rest
$100\,{\rm kpc}$ from the centre of the disc, but this time at an
angle of just $12^\circ$ from the disc plane. The choice of orbit is
motivated by a satellite in one of our cosmological simulations. The
parameters for the disc, halo, and subhalo are the same as in
\S\ref{ssec:toy_model_1}.

In the upper panel of Fig. \ref{fig:toy_model_2_orbits} we plot the
$z$ coordinates of the position vector ${\bf r}_d$ of the disc and
${\bf r}_h$, the position vector of the peak halo density.  Evidently,
the halo density peak and centre of the disc closely track one
another. We next define a common position vector for the disc and
inner halo:
\begin{equation}
{\bf r}_{dh} = \frac{M_d {\bf r}_d + M_{h,20} {\bf r}_{h,20}}
{M_d + M_{h,20}}
\end{equation}
\noindent where we have arbitrarily used $20\,{\rm kpc}$ to define the
inner halo. Given the closeness of ${\bf r}_d$ and ${\bf r}_h$, the
choice of this definition shouldn't effect the results that follow.

Our working hypothesis is that during periods between perigalactic passages,
the subhalo and the combination of the disc and inner halo can be
treated as a two-body system. That is, ${\bf r}_{dh}$ and ${\bf
  r}_{sh}$, where the latter is the position vector for the peak of
the subhalo, satisfy the equation
\begin{equation}
\frac{M_{dh}\textbf{r}_{dh} + M_{sh} \textbf{r}_{sh}}{M_{dh} +
  M_{sh}} = \textbf{r}_{c,0} + \textbf{v}_{c,0} t
\end{equation}
where $M_{dh}$ and $M_{sh}$ are the effective disc-halo and subhalo masses.

To test our hypothesis, we maximize the likelihood function
\begin{equation}
\ln \mathcal{L} = -\frac{\chi^2}{2\sigma_d^2} - N\ln\sigma_d / \kpc
\end{equation}
over the parameters ${\bf r}_{c,0}$, ${\bf v}_{c,0}$, $M_{dh}$,
and $M_{sh}$. Here, $N$ is the number of snapshots, {$\sigma_d$ is the RMS difference between the model and the data, }and
\begin{equation}
\chi^2 = \left\vert
\frac{M_{dh}\textbf{r}_{dh} + M_{sh} \textbf{r}_{sh}}{M_{dh} + M_{sh}}
- (\textbf{r}_{c,0} + \textbf{v}_{c,0} t )\right\vert^2~.
\end{equation}
Maximization is done in two $1 \, {\rm Gyr}$  epochs centered on $0.5\,{\rm
  Gyr}$ and $2\,{\rm Gyr}$, respectively. These windows are chosen to
avoid the pericentric passages at $0.85\,{\rm Gyr}$ and $3\,{\rm
  Gyr}$, when the subhalo is losing the most mass when the system will
be poorly described by a two-body model. The center-of-mass motion
lines, $\textbf{r}_{c,0} + \textbf{v}_{c,0} t$ for the two epochs are
shown in Fig. \ref{fig:toy_model_2_orbits}. The inferred effective
mass ratios $M_{sh}:M_{dh}$ are 16:1 and 85:1, respectively.  Thus,
the effective mass of the subhalo has been reduced by a factor of
$\sim 5$, presumably due to tidal stripping. We also find that
$\sigma_d= 190\,{\rm pc}$ and $290\,{\rm pc}$ for the two epochs.  The
fact that the values for $\sigma_d$ are as small as they are supports
our hypothesis of a two-body model for the disc-halo/subhalo system.

In Fig. \ref{fig:toy_model_2_mean_height} we plot the mean height
across the disc as a function of radius and time. The general
corrugation pattern is similar to the one seen in our first experiment
though the amplitude is smaller. The difference in amplitude is no doubt
due to the fact that we are averaging over azimuth in an experiment
that lacks azimuthal symmetry.

\section{Simulations} \label{sec:description}

In this section, we summarize our cosmological simulations beginning
with a brief description of our disc insertion scheme.

\begin{figure}
	\centering
	\includegraphics[width=0.85\textwidth]{../figures/rotation_curve.eps}
	\caption{Circular speed curves for cosmological models. The
          same disc is used in all runs, and its rotation curve
          contribution is shown by the black dashed line. The halo
          contributions for the Early and Warm subsuites are shown in
          the top panel as dot-dashed lines, while the total rotation
          curves are shown as solid curves. The bottom panel shows the
          same for the Late subsuite.} \label{fig:rotation_curves_iii}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.85\textwidth]{../figures/timescale.eps}\caption{The
          torque timescale, $T_\tau$, at initialization as a function of $R$.
          Colours are the same as in
          Fig. 5 with bauer2018b in the upper left panel, A in the upper right,
          B in the lower left, and C in the lower right.
          Line types are Early (solid), Warm (dot-dashed), and
          Late (dashed).} \label{fig:ratio_freqs}
\end{figure}

\begin{figure}
	%\centering
	\hspace{-0.75in}
	\includegraphics[width=1.2\textwidth]{../figures/substructure_r_vs_t.eps}
	\caption{Orbits of the most massive subhalo (solid curve)
          and second most massive subhalo (dashed curve) for the
          fiducial suite (top panel) and late insertion suite
          (bottom). These are the substructures listed as the top two
          entries for each simulation in Table
          \ref{tab:substructure}. The red line denotes $20\, \kpch =
          29.5 \, \kpc$. } \label{fig:substructure_orbits}
\end{figure}

\subsection{Disc Insertion Technique} \label{ssec:disc_insertio}

Our disc insertion scheme \citep{bauer2018a} expands upon the method
introduced by \citet{BerentzenShlosmanStellarDisks},
\citet{debuhr_2012}, and \citet{ys_2015}. It uses an iterative
procedure to initialize an axisymmetric, smooth disc in a cosmological
halo, that is generally clumpy and triaxial. The first step is to run
a pure dark matter simulation and identify a suitable halo. We then
rerun the simulation with a rigid disc potential that grows from zero
to that of the fully formed disc between an initial scale factor,
$a_g$, to a final scale factor, $a_l$.  When the simulation reaches
$a_l$, the rigid disc is replaced by an N-body system and the ``live''
disc-halo system is evolved to the present epoch at $a=1$.

For our pure dark matter simulation, we implement the zoom-in
technique of \citet{KatzQuasarZoom} and \citet{NavarroWhiteZoom}.  We
select Milky Way-like halos with no major mergers and use the results
of \cite{onorbe_etal_2014} to ensure minimal contamination of
low-resolution particles.  Our choice of cosmological parameters is
based on the results from Planck 2013 \citep{planck_2014}:
$H_0=67.9\,{\rm km\,s}^{-1}\,{\rm kpc}^{-1}$, $\Omega_b = 0.0481$,
$\Omega_0 = 0.306$, $\Omega_\Lambda = 0.694$, $\sigma_8 = 0.827$, and
$n_s = 0.962$. N-body initial conditions for the dark matter particles
are generated with the \textsc{music} code \citep{music}.

During its growth phase from $a_g=0.25$ to $a_l=0.5$\footnote{For
  reference, $a=0.25$ is 2.2 Gyr after the Big Bang, $a=0.5$ is 5.9
  Gyr after the Big Bang, and $a=0.75$ is 10.0 Gyr after the Big
  Bang. The age of this Universe at present day is 13.8 Gyr.}, the
disc is treated as a rigid body whose orientation and center-of-mass
position evolve according to the standard equations of rigid body
dynamics. \citet{bauer2018a} found that the rigid disc assumption is
actually quite true to the orientation and center of mass evolution of
a pure stellar disc. At $a_l$, we swap a live disc for the rigid one
using a modified version of \textsc{GalactICS}
\citep{KGGalactICSReference,WPDGalactICSReference}, which generates a
three-integral DF disc in the best axisymmetric approximation to the
halo \citet{bauer2018a}.

\subsection{Numerical Experiments} \label{ssec:numerical_experiments}

We present a concise description of our cosmological models in Table
\ref{tab:sims_iii}. The haloes are chosen such that the total circular
speed curves are similar to the Milky Way's. Note that the bauer2018b
simulations use a halo from \cite{bauer2018b}. The halos used in the
A, B, and C simulations are previously unpublished.  In the
simulations denoted as X.Early, the disc is inserted at $a_l=0.5$; for
those denoted X.Lake, $a_l=0.75$.  One of the benefits of disc
insertion techniques is the ability to modify disc properties in the
same halo. We exploit this capability in our third suite, denoted as
X.Warm, where we increase the disc's dynamical temperature as
measured by the Toomre $Q$ parameter \citep{toomre_q}.

Fig. \ref{fig:rotation_curves_iii} shows the circular speed curves for all
disc-halo combinations. There are relatively small variations in the
circular speed for $R<5\,{\rm kpc}$ where the disc dominates. The
variations are significant ($\sim 10\,{\rm km\,s}^{-1}$) near the peak
of the circular speed curve and even larger beyond the disc where the
halo dominates the gravitational potential. In all cases, the disc is
submaximal. In particular, the ratio of the square of the disc
contribution to the rotation curve and square of the total rotation
curve measured at $2.2\, R_d$, $\left. V_d^2/V_{tot}^2 \right
\vert_{R=2.2\,R_d}$, ranges between 0.37 and 0.46. These values are
well below the lower bound for a maximal disc advocated by
\citet{sackett_1997}.

In the discussion that follow, we focus on three archetypal
simulations:
\begin{enumerate}
\item \textbf{strong disc-subhalo interaction:} In B.Late the disc has close
  interactions with two subhaloes but no substantial disc-halo
  misalignment. One of these subhaloes passes twice through the plane
  of the disc.
\item \textbf{disc-halo misalignment:} C.Early presents a case a
  substantial misalignment from its host halo (see below) but
  relatively modest subhalo interactions with the disc.
\item \textbf{distant massive subhaloes:} The disc in A.Early is relatively
  well aligned to its host halo, but its most massive substructures do not
  interact closely with the disc.
\end{enumerate}

To illustrate how the simulations differ, consider
Fig.\,\ref{fig:torque} where we plot the component of the torque
perpendicular to the spin axis of the disc across the disc plane. In
B.late, we have the localized effect of a subhalo that passes through
the plane of the disc. On the other hand, in C.Early, we have a strong
$m=2$ torque due to the large-scale structure of the halo and
misalignment of the disc with any of the halo's approximate symmetry
axes.

\begin{figure}
	\centering
	\subfloat[]{\includegraphics[width=0.8\textwidth]{../figures/b_late_torque_a_0_885.eps}}\\
	\subfloat[]{\includegraphics[width=0.8\textwidth]{../figures/c_early_torque_a_0_5.eps}}
	\caption{The component of the torque perpendicular to the spin
          axis of the disc across the disc plane. Top
          panel is for B.Late at $a=0.885$ when the most massive
          subhalo is at pericentre.  The bottom panel is for C.Early
          at initialization ($a=0.5$). The circles correspond to $R_p
          = 2.2\,R_d=8.1\,\kpc$ and
          $20\,\kpch=29.5\,\kpc$.}\label{fig:torque}
\end{figure}

\begin{sidewaystable}
\centering
\caption{Table of simulations run in the main experiment. $M_h$ is the
  virial mass of the host halo, $R_h$ is the NFW scale length of the
  host halo, $c$ is the NFW concentration parameter of the host halo,
  $a_l$ is the live disc insertion scale factor, $\delta
  \Theta_{h,20}$ is the halo misalignment angle for the 20 kpc shell,
  $N_d$ is the number of disc particles, $Q$ is the Toomre stability
  parameter of the disc at initialization, and $f_{kud}$ is the
  fraction of disc mass that gets classified as kicked-out-disk (KOD)
  stars measured at present day with $\eta=8$.} \label{tab:sims_iii}
\begin{tabular}{l l l l l l l l l l}
\hline
       & $M_h\,\,(\solarm)$ & $R_h\,\,(\kpc)$ & $c$ & $a_l$ &  $N_d$ & $Q$ & $\cos\delta \Theta_{h,20}$ & $f_{kud}$\\
\hline
bauer2018.Early   & $1.0 \times 10^{12}$ & 15.9 &  9.84 &   0.50 & $1.0 \times 10^6$ & 1.1 & 0.16 & $1.2 \times 10^{-2}$\\
bauer2018.Warm    & $1.0 \times 10^{12}$ & 15.9 &  9.84 &   0.50 & $1.0 \times 10^6$ & 1.1 & 0.16 & $1.1 \times 10^{-2}$\\
bauer2018.Late    & $1.2 \times 10^{12}$ & 15.2 &  15.4 &   0.75 & $1.0 \times 10^6$ & 1.8 & 0.93 &$4.2 \times 10^{-3}$\\
A.Early           & $9.4 \times 10^{11}$ & 17.8 &  8.56 &   0.50 & $3.5 \times 10^6$ & 1.3 & 0.97 &$3.7 \times 10^{-3}$\\
A.Warm            & $9.4 \times 10^{11}$ & 17.8 &  8.56 &   0.50 & $3.5 \times 10^6$ & 1.8 & 0.97 & $1.6 \times 10^{-3}$\\
A.Late            & $1.1 \times 10^{12}$ & 18.5 &  12.1 &   0.75 & $3.5 \times 10^6$ & 1.3 & 0.98 & $6.8 \times 10^{-5}$\\
B.Early           & $5.9 \times 10^{11}$ & 10.4 &  12.6 &   0.50 & $3.5 \times 10^6$ & 1.3 & 0.91 & $8.5 \times 10^{-4}$\\
B.Warm            & $5.9 \times 10^{11}$ & 10.4 &  12.6 &   0.50 & $3.5 \times 10^6$ & 1.8 & 0.91 & $1.2 \times 10^{-3}$\\
B.Late            & $1.1 \times 10^{12}$ & 33.2 &  6.70 &   0.75 & $3.5 \times 10^6$ & 1.3 & 0.99 & $9.2 \times 10^{-4}$\\
C.Early           & $4.3 \times 10^{11}$ & 8.33 &  14.1 &   0.50 & $3.5 \times 10^6$ & 1.3 & 0.84 & $5.3 \times 10^{-3}$\\
C.Warm            & $4.3 \times 10^{11}$ & 8.33 &  14.1 &   0.50 & $3.5 \times 10^6$ & 1.8 & 0.84 & $5.6 \times 10^{-3}$\\
C.Late            & $6.7 \times 10^{11}$ & 13.6 &  14.0 &   0.75 & $3.5 \times 10^6$ & 1.3 & 0.92 & $1.0 \times 10^{-4}$\\
\hline
\end{tabular}

\end{sidewaystable}

\subsection{Halo Environments} \label{ssec:halo_env}

To further quantify the effect of the halo environment we define a
torque timescale:
\begin{equation}
T_\tau(R) = \frac{L(R)}{\tau_\perp(R)},
\end{equation}
where $L(R)$ is the magnitude of the angular momentum in a ring
centered at radius $R$ and $\tau_\perp(R)$ is the azimuthally-averaged
component of the torque that is perpendicular to the spin axis of the
disc.  The greater the misalignment between disc and halo, the larger
the torque of the halo on the disc, and the shorter the timescale
$T_\tau$. In Fig. \ref{fig:ratio_freqs}, we plot $T_\tau$ as a
function of radius at initialization for our models. In general, the
effect of tidal torques from the halo are smaller at late times when
much of the substructure and triaxiality of the halo has been washed
out. We also infer that the effect of the halo in torquing the disc is
largest in C.Early.

In Table \ref{tab:substructure}, we list the five most massive (at
initialization) substructures for each of our simulations while the
orbits of the two most massive substructures are shown in
Fig. \ref{fig:substructure_orbits}.  The orbits present a wide variety
of behaviour. For example, the subhaloes in bauer2018b.Early spend
most of their lives in the outer halo on roughly tangential orbits. By
contrast, the second most mass subhalo in C.late orbits to within
$\sim 20\,{\rm kpc}$ of the disc centre. The subhalos in A.early are
more massive but never reach to within $30\,{\rm kpc}$.

\begin{sidewaystable}
\caption{Table of the five most massive substructures at $a_l$ in all
  of the simulations detailed in Table \ref{tab:sims_iii}. The virial
  mass, $M_s$, and NFW scale radius, $R_s$, are
  given.}\label{tab:substructure}
	\begin{tabular}{|l l l l | l l l l|}
	\hline
	Subhalo Name ($z=1$) & $M_s \,(\solarm)$ & $R_s \, (\kpc)$ & $c$ & Subhalo Name ($z=0.333$) & $M_s \,(\solarm)$ & $R_s \, (\kpc)$ & $c$\\	
	\hline
    bauer2018b.Early.a & $1.4 \times 10^{10}$ & 3.75 & 9.92 & bauer2018b.Late.a & $1.6 \times 10^{10}$ & 3.35 & 16.5 \\
    bauer2018b.Early.b & $1.0 \times 10^{10}$ & 1.02 & 32.8 & bauer2018b.Late.e & $0.7 \times 10^{10}$ & 0.27 & 154. \\
    bauer2018b.Early.c & $0.5 \times 10^{10}$ & 2.60 & 10.1 & bauer2018b.Late.c & $0.3 \times 10^{10}$ & 0.07 & 485. \\
    bauer2018b.Early.d & $0.4 \times 10^{10}$ & 1.51 & 15.7 & bauer2018b.Late.d & $0.3 \times 10^{10}$ & 1.06 & 29.8 \\
    bauer2018b.Early.e & $0.2 \times 10^{10}$ & 0.28 & 69.0 & bauer2018b.Late.e & $0.2 \times 10^{10}$ & 0.38 &  76.7\\
	\hline
	A.Early.a & $1.8 \times 10^{10}$ & 2.69 & 15.3 & A.Late.a & $1.3 \times 10^{10}$ & 2.62 & 19.6 \\
	A.Early.b & $1.2 \times 10^{10}$ & 1.26 & 28.4 & A.Late.b & $0.9 \times 10^{10}$ & 0.61 & 74.8 \\
	A.Early.c & $1.0 \times 10^{10}$ & 0.71 & 47.9 & A.Late.c & $0.5 \times 10^{10}$ & 0.77 & 46.4 \\
	A.Early.d & $0.7 \times 10^{10}$ & 2.29 & 13.0 & A.Late.d & $0.4 \times 10^{10}$ & 1.69 & 19.7 \\
	A.Early.e & $0.5 \times 10^{10}$ & 1.73 & 13.2 & A.Late.e & $0.2 \times 10^{10}$ & 1.25 & 21.5 \\
	\hline
	B.Early.a & $1.1 \times 10^{10}$ & 1.76 & 19.4 & B.Late.a & $2.7 \times 10^{10}$ & 1.70 & 38.1 \\
	B.Early.b & $0.2 \times 10^{10}$ & 0.41 & 47.5 & B.Late.b & $1.2 \times 10^{10}$ & 2.27 & 22.0 \\
	B.Early.c & $0.2 \times 10^{10}$ & 0.67 & 28.4 & B.Late.c & $0.6 \times 10^{10}$ & 2.67 & 14.4 \\
	B.Early.d & $0.2 \times 10^{10}$ & 1.81 & 10.3 & B.Late.d & $0.4 \times 10^{10}$ & 1.42 & 24.0 \\
	B.Early.e & $0.2 \times 10^{10}$ & 0.54 & 34.0 & B.Late.e & $0.4 \times 10^{10}$ & 1.13 & 29.7 \\
	\hline
    C.Early.a & $1.0 \times 10^{10}$ & 1.78 & 18.5 & C.Late.a & $0.5 \times 10^{10}$ & 0.18 & 194. \\
	C.Early.b & $0.6 \times 10^{10}$ & 0.92 & 30.0 & C.Late.b & $0.4 \times 10^{10}$ & 6.37 & 5.46 \\
	C.Early.c & $0.3 \times 10^{10}$ & 2.24 & 10.2 & C.Late.c & $0.4 \times 10^{10}$ & 0.75 & 46.2 \\
	C.Early.d & $0.1 \times 10^{10}$ & 0.91 & 18.7 & C.Late.d & $0.3 \times 10^{10}$ & 1.26 & 26.1 \\
	C.Early.e & $0.1 \times 10^{10}$ & 0.25 & 60.3 & C.Late.e & $0.3 \times 10^{10}$ & 6.78 & 4.81 \\
	\hline
	\end{tabular}
\end{sidewaystable}

\section{Structure and Evolution of Simulation Models} \label{sec:evolution}

In this section we summarize the evolution of the models in our
simulations.  We begin with a discussion of bar formation and then
turn to bending and warping of the discs.

\subsection{Bar Formation}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{../figures/surface_density_ratios.eps} \caption{The
      ratio of surface density, $\Sigma$, to azimuthally averaged
      surface density, $\Sigma_0$ for the C.Early, C.Warm, and C.Late
      models. The circles correspond to $R_p = 2.2\, R_d = 8.1 \,\kpc$
      and $20 \,\kpch = 29.5 \, \kpc$.}
	\label{fig:final_densities}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.85\textwidth]{../figures/a_2_all_models_four_panel.eps}
	\caption{Bar strength as measured by $A_2(<R_p)$ as a function of time for all models.} \label{fig:a_2}
\end{figure}
